---
title: "Optimize activation functions"
author: "George Makarov"
date: "25/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo    = TRUE)
knitr::opts_chunk$set(warning = TRUE)
```

## Optimize activation functions

Some packages use `ifelse()` in activation functions. This function is a
vectorized version of the standard control combination of functions
`if (condition) <X> else <Y>`. Function `ifelse()` is more programmer efficient.
One can understand it easier too. However it is **less* computationally efficient
than its hardcoded alternative.

The example below compares `ifelse()` to alternative approach. We start from
preparing a large matrix.

```{r}
n  <- c(-1, 0, 1, 2, 3, 4, 5)
s  <- 3e7
n  <- sample(x = n, size = s, replace = T)
w  <- matrix(n, nrow = 5, ncol = s / 5)
dim(w)
```

Lets make two `relu` functions to demonstrate the difference.

```{r}
relu_ifelse <- function(x){return(ifelse(x >= 0, x, 0))}
relu_hard   <- function(x){out <- x; out[out < 0] <- 0; return(out)}
```

Now lets compare the execution time of each function.

```{r}
system.time({t1 <- relu_ifelse(w)})
system.time({t2 <- relu_hard(w)})
```

Check tht both results are identical:

```{r}
identical(t1, t2)
```

## Summary

Using hardcoded alternative in activation function is **4** times faster than
the vectorized `ifelse()` function.
