---
title: "Regularization cost"
author: "Georgy Makarov"
date: "2/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

This document justifies the use of `eval` function in `compute_cost_reg` function.
It shows that `eval` is not slower than hard coded solution. At the same time
`eval` allows us to find sum of costs for any number of layers without massive
code constructions.

## Initial data

We reproduce a set of weight matrices as it appears after we have initialized
the parameters of *nerual network*. Both number of rows and number of columns 
reflect the sizes of layers of our network. Values of the matrices are random.
We specifically make the matrices large enough to see the difference between
computations.

```{r}
w1 <- matrix(rnorm(160000, 5, 2.3), nrow = 400, ncol = 400)
w2 <- matrix(rnorm(80000, 3, 1.3),  nrow = 200, ncol = 400)
w3 <- matrix(rnorm(20000, 1, 0.3),  nrow = 100, ncol = 200)
```


## Comparison

There is no difference in execution time between hard coded and `eval` function.
This allows us to use the latter for our computations.

```{r}
system.time(sum(w1 ^ 2) + sum(w2 ^ 2) + sum(w3 ^ 2))
```

```{r}
system.time(eval(parse(text = "sum(w1 ^ 2) + sum(w2 ^ 2) + sum(w3 ^ 2)")))
```
